---
title: "Impact of Geographic Location and Listing Features on Airbnb Guest Satisfaction"
author: "Group 8: Yilin Xiao,Violet Yong, Yuqi Wang, Victor Zhan"
date: "2025-04-22"
output: html_document
---

# Research Question 1:

**Does the geographic location of Airbnb listings impact customer satisfaction?**

- **Null Hypothesis (H₀):** Geographic location has no significant impact.
- **Alternative Hypothesis (H₁):** Geographic location significantly impacts customer satisfaction.

---

## Load Libraries & Data

```{r}
# Load required libraries
library(tidyverse)

# Load cleaned dataset
df <- read_csv("~/Downloads/listings_cleaned-2.csv")
```

---

## 1. K-Means Clustering on Latitude & Longitude

```{r}
# Apply K-means clustering
set.seed(123)
coords <- df %>% select(latitude, longitude)
kmeans_result <- kmeans(coords, centers = 5)
df$geo_cluster <- as.factor(kmeans_result$cluster)

# Boxplot of review scores by cluster
boxplot(review_scores_rating ~ geo_cluster, data = df,
        main = "Review Scores by Geo Cluster",
        xlab = "Geo Cluster", ylab = "Review Score", col = "skyblue")
```

---

## 2. One-Way ANOVA by Neighborhood

```{r}
anova_result <- aov(review_scores_rating ~ neighbourhood_cleansed, data = df)

cat("### ANOVA Result\n")
summary(anova_result)
```

---

## 3. Kruskal-Wallis Test (Non-parametric)

```{r}
kruskal_result <- kruskal.test(review_scores_rating ~ neighbourhood_cleansed, data = df)

cat("### Kruskal-Wallis Result\n")
print(kruskal_result)
```

---

## 4. Spatial Visualization of Review Scores

```{r}
ggplot(df, aes(x = longitude, y = latitude, color = review_scores_rating)) +
  geom_point(alpha = 0.6) +
  scale_color_viridis_c() +
  labs(title = "Spatial Distribution of Review Scores",
       x = "Longitude", y = "Latitude") +
  theme_minimal()
```

---

## 5. Principal Component Analysis

```{r}
geo_pca <- prcomp(df %>% select(latitude, longitude), scale. = TRUE)

cat("### PCA Summary\n")
summary(geo_pca)

biplot(geo_pca)
```

---

## Conclusion

All methods — K-means clustering, ANOVA, Kruskal-Wallis test, and spatial visualizations — suggest that **geographic location significantly impacts guest satisfaction** on Airbnb. Listings in different neighborhoods or spatial clusters receive systematically different review scores, allowing us to **reject the null hypothesis** with high confidence.



# Research Question 2:

**How do listing features influence review score ratings, and what areas could be targeted for improvement?**

- **Null Hypothesis (H₀):** Listing features do not significantly influence review score ratings.
- **Alternative Hypothesis (H₁):** Listing features significantly influence review score ratings, indicating areas for potential improvement.

---

## Load Libraries & Data

```{r}
# Install and load necessary packages
# install.packages("tidyverse")
# install.packages("car")
library(tidyverse)
library(car)

# Load cleaned dataset

df <- read_csv("~/Downloads/listings_cleaned-2.csv")
```

---

## 1. Multiple Linear Regression (Simplified)

```{r}
# Fit a simplified model to avoid multicollinearity
model_fixed <- lm(review_scores_rating ~ price + bedrooms + bathrooms_num +
                    room_type + number_of_reviews + host_is_superhost,
                  data = df)

# Model summary
summary(model_fixed)
```

---

## 2. Check for Multicollinearity (VIF)

```{r}
# Check VIF (Variance Inflation Factor)
vif(model_fixed)
```

---

## 3. Visualizing Effects of Key Features

```{r}
# Price vs Review Score
ggplot(df, aes(x = price, y = review_scores_rating)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", color = "blue") +
  labs(title = "Price vs Review Score", x = "Price", y = "Review Score") +
  theme_minimal()

# Review Score by Superhost Status
ggplot(df, aes(x = host_is_superhost, y = review_scores_rating)) +
  geom_boxplot(fill = "lightgreen") +
  labs(title = "Review Scores by Superhost Status", x = "Superhost", y = "Review Score") +
  theme_minimal()
```

---

## 4.PCA on Numeric Listing Features

```{r}
# Select and scale numeric features
feature_data <- df %>% select(price, bedrooms, bathrooms_num, number_of_reviews, minimum_nights) %>% na.omit()
pca_features <- prcomp(feature_data, scale. = TRUE)

# Summary of PCA
summary(pca_features)

# Biplot
biplot(pca_features)
```

---

## Conclusion

The revised regression model shows that key listing features — such as price, room type, superhost status, and number of bedrooms/bathrooms — have a meaningful impact on review scores. Multicollinearity has been addressed, and visualizations support these findings. Therefore, we reject the null hypothesis and conclude that **listing features significantly influence guest satisfaction**, helping identify **areas for improvement**.


# Conclusion and Recommendations

## Figure 8

```{r heatmap-borough-review-scores, message=FALSE, warning=FALSE}
# Load required packages
library(dplyr)
library(ggplot2)
library(tidyr)

# Load your dataset (make sure the file is in your working directory)
listings <- read.csv("~/Downloads/listings_cleaned-2.csv")

# Extract review score columns
score_cols <- grep("^review_scores_", names(listings), value = TRUE)

# Group by borough and calculate average scores for each category
borough_scores <- listings %>%
  group_by(neighbourhood_group_cleansed) %>%
  summarise(across(all_of(score_cols), ~ mean(.x, na.rm = TRUE))) %>%
  pivot_longer(cols = all_of(score_cols), names_to = "score_type", values_to = "avg_score")

# Clean column labels for display
borough_scores$score_type <- gsub("review_scores_", "", borough_scores$score_type)

# Create the red/blue heatmap
ggplot(borough_scores, aes(x = score_type, y = neighbourhood_group_cleansed, fill = avg_score)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "steelblue", mid = "white", high = "red", midpoint = 4.7, na.value = "gray90") +
  labs(
    title = "Average Review Scores by Borough and Category",
    x = "Review Score Category",
    y = "Borough",
    fill = "Avg Score"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


## Figure 9

```{r rq2-coefficient-plot, message=FALSE, warning=FALSE} 
# Load packages
library(ggplot2)
library(broom)

# Run the regression model (if not already done)
model <- lm(review_scores_rating ~ price + bedrooms + bathrooms_num + 
              room_type + number_of_reviews + host_is_superhost, data = listings)

# Extract coefficients and clean up terms
coef_df <- tidy(model) %>%
  filter(term != "(Intercept)") %>%
  mutate(term = case_when(
    term == "price" ~ "Price (USD)",
    term == "bedrooms" ~ "Number of Bedrooms",
    term == "bathrooms_num" ~ "Number of Bathrooms",
    term == "number_of_reviews" ~ "Number of Reviews",
    term == "host_is_superhost" ~ "Superhost Status",
    grepl("room_type", term) ~ gsub("room_type", "Room Type: ", term),
    TRUE ~ term
  ))

# Plot the coefficients with 95% CI
ggplot(coef_df, aes(x = reorder(term, estimate), y = estimate)) +
  geom_point(color = "darkblue", size = 3) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2, color = "gray40") +
  coord_flip() +
  labs(
    title = "Effect of Listing Features on Review Scores",
    x = "Listing Feature",
    y = "Effect on Review Score"
  ) +
  theme_minimal(base_size = 12)

```


## Figure 10

```{r feature-reviewscore-heatmap, message=FALSE, warning=FALSE}
# Load required libraries
library(dplyr)
library(tidyr)
library(ggplot2)

# Identify review score columns
score_cols <- grep("^review_scores_", names(df), value = TRUE)

# Combine listing features into a single label (e.g., "Private room, Superhost: TRUE")
df <- df %>%
  mutate(feature_group = paste0(room_type, ", Superhost: ", host_is_superhost))

# Calculate average review scores for each feature group
feature_scores <- df %>%
  group_by(feature_group) %>%
  summarise(across(all_of(score_cols), ~ mean(.x, na.rm = TRUE))) %>%
  pivot_longer(cols = all_of(score_cols), names_to = "review_type", values_to = "avg_score")

# Clean up review score category labels
feature_scores$review_type <- gsub("review_scores_", "", feature_scores$review_type)

# Plot the heatmap
ggplot(feature_scores, aes(x = review_type, y = feature_group, fill = avg_score)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "steelblue", mid = "white", high = "red", midpoint = 4.7) +
  labs(
    title = "Average Review Scores by Listing Feature and Review Category",
    x = "Review Category",
    y = "Listing Feature Group",
    fill = "Avg Score"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

